Natural Stories model analyses

### 2021-08-23 Analysis/models.Rmd
Roger completed bootstrapping work for models, both Maze and SPR

```{r bootstrap code}
require(OpenImageR)
gam_boot <- function(dat,formula,data_frames_for_prediction,N=100,method="REML",...) {
  models <- lapply(1:N,function(seed) {
    set.seed(seed)
    #dat[[weights_name]] <- tabulate(sample(1:nrow(dat),floor(nrow(dat)/5),replace=TRUE),nbins=nrow(dat))
    dat$mgcv_weights <- tabulate(sample(1:nrow(dat),nrow(dat),replace=TRUE),nbins=nrow(dat))
    #print(head(dat$mgcv_weights))
    #print(head(dat[[weights_name]]))
    #m <- gam(formula,data=dat,weights=dat[[weights_name]],...)
    m <- gam(formula,data=dat,weights=mgcv_weights,method=method,...)
    #print(m$deviance)
    return(m)
  })
  results <- lapply(names(data_frames_for_prediction),function(term) {
    predictions <-sapply(models,function(m)
      predict(m,newdata=data_frames_for_prediction[[term]],type="terms",terms=term))
    #print(dim(predictions))
    return(predictions)
  })
  #print(length(results))
  names(results) <- names(data_frames_for_prediction)
  return(results)
}

summarize_gam_boot_results <- function(gam_boot_results,data_frames_for_prediction,term_names,model_name,point_estimation_function=mean,quantiles=c(0.025,0.975)) {
  result <- lapply(1:length(data_frames_for_prediction),function(i) {
    #print(term_names[[i]])
    res <- tibble(x=data_frames_for_prediction[[i]][[term_names[[i]]]],
                  rt=apply(gam_boot_results[[i]],1,point_estimation_function),
                  CI_lower=apply(gam_boot_results[[i]],1,function(x) quantile(x,0.1)),
                  CI_upper=apply(gam_boot_results[[i]],1,function(x) quantile(x,0.9)),
                  model=model_name,
                  term=term_names[[i]])
    names(res)[names(res)=="x"] <- term_names[[i]]
    return(res)
  })
  return(result)
}

@@

write_rds(all_surprisal_predictions,"bootstrapped_GAM_surprisal_predictions.rds")
```

@@

@@

gam1 <- read_rds(here("Analysis/models/spr_gam_predictions.rds")) %>%  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
  geom_line()+
  geom_ribbon(alpha=.3)+
  facet_grid(s~model)+
  coord_cartesian(xlim=c(0,28))+
  labs(x="Surprisal (bits)", y="Reaction Time (ms)")+theme(axis.ticks.x=element_blank(), axis.title.x=element_blank(), axis.text.x=element_blank(), plot.margin=margin(t=0,r=0,b=0,l=0,unit="pt"))

>>>>

```{r maze bootstrap code}
require(OpenImageR)
gam_boot <- function(dat,formula,data_frames_for_prediction,N=100,method="REML",...) {
  models <- lapply(1:N,function(seed) {
    set.seed(seed)
    #dat[[weights_name]] <- tabulate(sample(1:nrow(dat),floor(nrow(dat)/5),replace=TRUE),nbins=nrow(dat))
    dat$mgcv_weights <- tabulate(sample(1:nrow(dat),nrow(dat),replace=TRUE),nbins=nrow(dat))
    #print(head(dat$mgcv_weights))
    #print(head(dat[[weights_name]]))
    #m <- gam(formula,data=dat,weights=dat[[weights_name]],...)
    m <- gam(formula,data=dat,weights=mgcv_weights,method=method,...)
    #print(m$deviance)
    return(m)
  })
  results <- lapply(names(data_frames_for_prediction),function(term) {
    predictions <-sapply(models,function(m)
      predict(m,newdata=data_frames_for_prediction[[term]],type="terms",terms=term))
    #print(dim(predictions))
    return(predictions)
  })
  #print(length(results))
  names(results) <- names(data_frames_for_prediction)
  return(results)
}

summarize_gam_boot_results <- function(gam_boot_results,data_frames_for_prediction,term_names,model_name,point_estimation_function=mean,quantiles=c(0.025,0.975)) {
  result <- lapply(1:length(data_frames_for_prediction),function(i) {
    #print(cat(i,"\t",term_names[[i]]))
    res <- tibble(x=data_frames_for_prediction[[i]][[term_names[[i]]]],
                  rt=apply(gam_boot_results[[i]],1,point_estimation_function),
                  CI_lower=apply(gam_boot_results[[i]],1,function(x) quantile(x,quantiles[1])),
                  CI_upper=apply(gam_boot_results[[i]],1,function(x) quantile(x,quantiles[2])),
                  model=model_name,
                  term=term_names[[i]]
                  )
    names(res)[names(res)=="x"] <- term_names[[i]]
    return(res)
  })
  return(result)
}

@@

write_rds(all_bootstrapped_predictions,"bootstrapped_GAM_predictions.rds")
write_rds(all_surprisal_predictions,"bootstrapped_GAM_surprisal_predictions.rds")
```

@@

```{r spr bootstrapping}
surprisal_range <- seq(0,25,by=0.01)
freq_range <- seq(-9.5,7,by=0.01)
len_range <- seq(-3.4,10.6,by=0.01)
surprisal_terms <- c("surp","past_surp","past2_surp","past3_surp")
freq_terms <- c("freq_center","past_freq_center","past2_freq_center","past3_freq_center")
len_terms <- c("length_center","past_length_center","past2_length_center","past3_length_center")

dat_for_surp_prediction <- expand_grid(surp=seq(0,25,by=0.01),freq_center=0,length_center=0,past_surp=0,past_freq_center=0,past_length_center=0,past2_surp=0,past2_freq_center=0,past2_length_center=0,past3_surp=0,past3_freq_center=0,past3_length_center=0)
dat_for_freq_prediction <- expand_grid(surp=0,freq_center=seq(-9.5,7,by=0.01),length_center=0,past_surp=0,past_freq_center=0,past_length_center=0,past2_surp=0,past2_freq_center=0,past2_length_center=0,past3_surp=0,past3_freq_center=0,past3_length_center=0)
dat_for_length_prediction <- expand_grid(surp=0,freq_center=0,length_center=seq(-3.4,10.6,by=0.01),past_surp=0,past_freq_center=0,past_length_center=0,past2_surp=0,past2_freq_center=0,past2_length_center=0,past3_surp=0,past3_freq_center=0,past3_length_center=0)
dat_for_past_surp_prediction <- expand_grid(surp=0,freq_center=0,length_center=0,past_surp=seq(0,25,by=0.01),past_freq_center=0,past_length_center=0,past2_surp=0,past2_freq_center=0,past2_length_center=0,past3_surp=0,past3_freq_center=0,past3_length_center=0)
dat_for_past_freq_prediction <- expand_grid(surp=0,freq_center=0,length_center=0,past_surp=0,past_freq_center=seq(-9.5,7,by=0.01),past_length_center=0,past2_surp=0,past2_freq_center=0,past2_length_center=0,past3_surp=0,past3_freq_center=0,past3_length_center=0)
dat_for_past_len_prediction <- expand_grid(surp=0,freq_center=0,length_center=0,past_surp=0,past_freq_center=0,past_length_center=seq(-3.4,10.6,by=0.01),past2_surp=0,past2_freq_center=0,past2_length_center=0,past3_surp=0,past3_freq_center=0,past3_length_center=0)
dat_for_past2_surp_prediction <- expand_grid(surp=0,freq_center=0,length_center=0,past_surp=0,past_freq_center=0,past_length_center=0,past2_surp=seq(0,25,by=0.01),past2_freq_center=0,past2_length_center=0,past3_surp=0,past3_freq_center=0,past3_length_center=0)
dat_for_past2_freq_prediction <- expand_grid(surp=0,freq_center=0,length_center=0,past_surp=0,past_freq_center=0,past_length_center=0,past2_surp=0,past2_freq_center=seq(-9.5,7,by=0.01),past2_length_center=0,past3_surp=0,past3_freq_center=0,past3_length_center=0)
dat_for_past2_len_prediction <- expand_grid(surp=0,freq_center=0,length_center=0,past_surp=0,past_freq_center=0,past_length_center=0,past2_surp=0,past2_freq_center=0,past2_length_center=seq(-3.4,10.6,by=0.01),past3_surp=0,past3_freq_center=0,past3_length_center=0)
dat_for_past3_surp_prediction <- expand_grid(surp=0,freq_center=0,length_center=0,past_surp=0,past_freq_center=0,past_length_center=0,past2_surp=0,past2_freq_center=0,past2_length_center=0,past3_surp=seq(0,25,by=0.01),past3_freq_center=0,past3_length_center=0)
dat_for_past3_freq_prediction <- expand_grid(surp=0,freq_center=0,length_center=0,past_surp=0,past_freq_center=0,past_length_center=0,past2_surp=0,past2_freq_center=0,past2_length_center=0,past3_surp=0,past3_freq_center=seq(-9.5,7,by=0.01),past3_length_center=0)
dat_for_past3_len_prediction <- expand_grid(surp=0,freq_center=0,length_center=0,past_surp=0,past_freq_center=0,past_length_center=0,past2_surp=0,past2_freq_center=0,past2_length_center=0,past3_surp=0,past3_freq_center=0,past3_length_center=seq(-3.4,10.6,by=0.01))

term_names <- c("surp","freq_center","length_center","past_surp","past_freq_center","past_length_center","past2_surp","past2_freq_center","past2_length_center","past3_surp","past3_freq_center","past3_length_center")
dats_for_prediction <- list(
  's(surp)'=dat_for_surp_prediction,
  'ti(freq_center)'=dat_for_freq_prediction,
  'ti(length_center)'=dat_for_length_prediction,
  's(past_surp)'=dat_for_past_surp_prediction,
  'ti(past_freq_center)'=dat_for_past_freq_prediction,
  'ti(past_length_center)'=dat_for_past_len_prediction,
  's(past2_surp)'=dat_for_past2_surp_prediction,
  'ti(past2_freq_center)'=dat_for_past2_freq_prediction,
  'ti(past2_length_center)'=dat_for_past2_len_prediction,
  's(past3_surp)'=dat_for_past3_surp_prediction,
  'ti(past3_freq_center)'=dat_for_past3_freq_prediction,
  'ti(past3_length_center)'=dat_for_past3_len_prediction)

N_replicates <- 101
res_ngram_spr <- gam_boot(ngram_mean_spr,spr_formula_interact,dats_for_prediction,N=N_replicates)
bootstrapped_gam_predictions_ngram_spr <- summarize_gam_boot_results(res_ngram_spr,dats_for_prediction,term_names,"5-gram")
res_grnn_spr <- gam_boot(grnn_mean_spr,spr_formula_interact,dats_for_prediction,N=N_replicates)
bootstrapped_gam_predictions_grnn_spr <- summarize_gam_boot_results(res_grnn_spr,dats_for_prediction,term_names,"GRNN")
res_txl_spr <- gam_boot(txl_mean_spr,spr_formula_interact,dats_for_prediction,N=N_replicates)
bootstrapped_gam_predictions_txl_spr <- summarize_gam_boot_results(res_txl_spr,dats_for_prediction,term_names,"Transformer-XL")
res_gpt_spr <- gam_boot(gpt_mean_spr,spr_formula_interact,dats_for_prediction,N=N_replicates)
bootstrapped_gam_predictions_gpt_spr <- summarize_gam_boot_results(res_gpt_spr,dats_for_prediction,term_names,"GPT-2")
all_bootstrapped_predictions_spr <- list(ngram=bootstrapped_gam_predictions_ngram_spr,
                                     grnn=bootstrapped_gam_predictions_grnn_spr,
                                     txl=bootstrapped_gam_predictions_txl_spr,
                                     gpt=bootstrapped_gam_predictions_gpt_spr)
write_rds(all_bootstrapped_predictions_spr,"bootstrapped_spr_GAM_predictions.rds")

term_conversions <- c("surp"="Current","past_surp"="Previous","past2_surp"="2Previous","past3_surp"="3Previous")
all_surprisal_predictions_spr <- rbind(rename(bootstrapped_gam_predictions_ngram_spr[[1]],surprisal=surp),
                                   rename(bootstrapped_gam_predictions_ngram_spr[[4]],surprisal=past_surp),
                                   rename(bootstrapped_gam_predictions_ngram_spr[[7]],surprisal=past2_surp),
                                   rename(bootstrapped_gam_predictions_ngram_spr[[10]],surprisal=past3_surp),
                                   rename(bootstrapped_gam_predictions_grnn_spr[[1]],surprisal=surp),
                                   rename(bootstrapped_gam_predictions_grnn_spr[[4]],surprisal=past_surp),
                                   rename(bootstrapped_gam_predictions_grnn_spr[[7]],surprisal=past2_surp),
                                   rename(bootstrapped_gam_predictions_grnn_spr[[10]],surprisal=past3_surp),
                                   rename(bootstrapped_gam_predictions_txl_spr[[1]],surprisal=surp),
                                   rename(bootstrapped_gam_predictions_txl_spr[[4]],surprisal=past_surp),
                                   rename(bootstrapped_gam_predictions_txl_spr[[7]],surprisal=past2_surp),
                                   rename(bootstrapped_gam_predictions_txl_spr[[10]],surprisal=past3_surp),
                                   rename(bootstrapped_gam_predictions_gpt_spr[[1]],surprisal=surp),
                                   rename(bootstrapped_gam_predictions_gpt_spr[[4]],surprisal=past_surp),
                                   rename(bootstrapped_gam_predictions_gpt_spr[[7]],surprisal=past2_surp),
                                   rename(bootstrapped_gam_predictions_gpt_spr[[10]],surprisal=past3_surp)
                                   ) %>%
  mutate(s=factor(term_conversions[term],levels=term_conversions)) %>%
  select(-term)
write_rds(all_surprisal_predictions_spr,"bootstrapped_spr_GAM_surprisal_predictions.rds")

@@

#gam1 <- read_rds(here("Analysis/models/spr_gam_predictions.rds")) %>%  
gam1 <- read_rds(here("Analysis/bootstrapped_spr_GAM_surprisal_predictions.rds")) %>%  
  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
  geom_line()+
  geom_ribbon(alpha=.3)+
  facet_grid(s~model)+
  coord_cartesian(xlim=c(0,28))+
  labs(x="Surprisal (bits)", y="Reaction Time (ms)")+theme(axis.ticks.x=element_blank(), axis.title.x=element_blank(), axis.text.x=element_blank(), plot.margin=margin(t=0,r=0,b=0,l=0,unit="pt"))
